---
title: "Adam Ercanbrack Unit2 LiveSession Assignment"
author: "Adam E."
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("gganimate")
```

```{r load_library}

library(tidyverse)
library(ggplot2)
library(ggthemes)
library(plotly)

```

# Data Import from csv
```{r csv_import2}
Path <- "D:/University/SMU/Doing_Data_Science/DDS_repository/Doing-Data-Science/Chapter2/PlayersBBall.csv"
BBallP = read.csv(Path, header=TRUE)

summary(BBallP)
# Keep only rows in the position column where the value is not NA or empty.
BBallP_C <- BBallP[!(is.na(BBallP$position) | BBallP$position ==""),]

# Verify that all empties are removed.
x <- unique(BBallP_C$position)
x

BBallP_C

summary(BBallP_C)


```
# Q:1 - Use the PlayerBBall.csv dataset to visually represent (summarize) 
the number of players in each position.
```{r Summary_by_position}
# visually summarize the players in each position
BBallP_C %>% 
  ggplot(mapping = aes(x = position)) + 
  geom_bar(color = "black", fill = "blue") +
  ggtitle("Summary of Basketball Players") + # Title
  xlab("Position") + ylab("Number of Players") + # x and y labels
  theme_economist()
```
# Q:2 - Use the dataset to visually investigate whether the distribution of the 
weight of centers (C) is greater than the distribution of the weight of forwards (F).
```{r Weight_distribution }
# Create a new DF that only hold the positions of C or F.
BBallC_F <- BBallP_C[(BBallP_C$position == "C" | BBallP_C$position == "F"),]
BBallC_F

# Replace the value names
BBallC_F2 <- BBallC_F %>%
mutate(position = recode(position, "C" = "Center", "F" = "Forward")) %>%
        arrange(weight, position )
BBallC_F2

# Use a boxplot to visualize the weights
BBallC_F2 %>% 
  ggplot(aes(x = weight, y = BBallC_F2$position, )) + geom_boxplot(color= "black", fill = "grey") +
  ggtitle("Weight Distribution: Forwards vs. Centers") + # Title
  xlab("Weight of Player") + ylab("Player's Position") + # x and y labels
  theme_economist()

```
# Q:3 - Use the dataset to visually investigate if the distribution of the 
height of centers (C) is greater than the distribution of the height of 
forwards (F).
```{r Height_distribution }
# Create a new DF that only hold the positions of C or F.
BBallC_F <- BBallP_C[(BBallP_C$position == "C" | BBallP_C$position == "F"),]
BBallC_F

# Replace the value names
BBallC_F3 <- BBallC_F %>%
mutate(position = recode(position, "C" = "Center", "F" = "Forward")) %>%
  mutate(height = recode(height, "6-0" = "72", "6-1" = "73", "6-2" = "74", 
                         "6-3" = "75", "6-4" = "76", "6-5" = "77", "6-6" = "78",
                         "6-7" = "79", "6-8" = "80", "6-9" = "81", "6-10" = "82",
                         "6-11" = "83", "7-0" = "84", "7-1" = "85", "7-2" = "86",
                         "7-3" = "87", "7-4" = "88", "7-5" = "89", "7-6" = "90", 
                         "7-7" = "91",)) %>%
        arrange(height, position )
# change the height column from chr to int
BBallC_F3$height <- as.numeric(BBallC_F3$height)
BBallC_F3

# Use a boxplot to visualize the heights
BBallC_F3 %>% 
  ggplot(aes(x = BBallC_F3$position, y = BBallC_F3$height)) + 
  geom_boxplot(color= "black", fill = "grey") +
  ggtitle("Height Distribution: Centers vs. Forwards") + # Title
  xlab("Player's Position") + ylab("Height (In inches)") + # x and y labels
  theme_economist()

```
# Q:4 - Use the dataset to visually investigate if the distribution of height 
is different between any of the positions.
```{r Height_distribution }
# Replace the value names
BBallP_C4 <- BBallP_C %>%

  mutate(height = recode(height, "6-0" = "72", "6-1" = "73", "6-2" = "74", 
                         "6-3" = "75", "6-4" = "76", "6-5" = "77", "6-6" = "78",
                         "6-7" = "79", "6-8" = "80", "6-9" = "81", "6-10" = "82",
                         "6-11" = "83", "7-0" = "84", "7-1" = "85", "7-2" = "86",
                         "7-3" = "87", "7-4" = "88", "7-5" = "89", "7-6" = "90", 
                         "7-7" = "91",)) %>%
        arrange(height, position )

# change the height column from chr to int
BBallP_C4 $height <- as.numeric(BBallP_C4$height)
BBallP_C4 

# Use a boxplot to visualize the heights
BBallP_C4  %>% 
  ggplot(aes(x = BBallP_C4$position, y = BBallP_C4 $height)) + 
  geom_boxplot(color= "black", fill = "grey") +
  ggtitle("Height Distribution: All Positions") + # Title
  xlab("Player's Position") + ylab("Height (In inches)") + # x and y labels
  theme_economist()
```
# Q:5 - Use the dataset to investigate how the player’s height is related to 
the player’s weight. How does height change as the weight changes?  
```{r Weight_Height_relate}
# Replace the value names
BBallP_C4 <- BBallP_C %>%
  mutate(height = recode(height, "6-0" = "72", "6-1" = "73", "6-2" = "74", 
                         "6-3" = "75", "6-4" = "76", "6-5" = "77", "6-6" = "78",
                         "6-7" = "79", "6-8" = "80", "6-9" = "81", "6-10" = "82",
                         "6-11" = "83", "7-0" = "84", "7-1" = "85", "7-2" = "86",
                         "7-3" = "87", "7-4" = "88", "7-5" = "89", "7-6" = "90", 
                         "7-7" = "91",)) %>%
        arrange(height, position )

# change the height column from chr to int
BBallP_C4 $height <- as.numeric(BBallP_C4$height)
BBallP_C4 

# Jitter plot with a line and labels in economist theme
BBallP_C4 %>% 
  ggplot(aes(x = height, y = weight)) + geom_jitter(aes(size=weight)) +
  geom_smooth(method=lm, se=FALSE) + # Add a regression line
  ggtitle("Relation of Weight to Height: All Positions") + # Title
  xlab("Height (In inches)") + ylab("Player's Weight") + # x and y labels
  theme_economist()

```
# Q:6 - Is their any difference in the relationship between height and weight 
between positions?  Are height and weight related differently for different 
positions.  
```{r Weight_Height_relate}
# Replace the value names
BBallP_C4 <- BBallP_C %>%
  mutate(height = recode(height, "6-0" = "72", "6-1" = "73", "6-2" = "74", 
                         "6-3" = "75", "6-4" = "76", "6-5" = "77", "6-6" = "78",
                         "6-7" = "79", "6-8" = "80", "6-9" = "81", "6-10" = "82",
                         "6-11" = "83", "7-0" = "84", "7-1" = "85", "7-2" = "86",
                         "7-3" = "87", "7-4" = "88", "7-5" = "89", "7-6" = "90", 
                         "7-7" = "91",)) %>%
        arrange(height, position )

# change the height column from chr to int
BBallP_C4 $height <- as.numeric(BBallP_C4$height)
BBallP_C4 

# Jitter plot with a line and labels in economist theme
BBallP_C4 %>% 
  ggplot(aes(x = height, y = weight, color = position)) + 
  geom_jitter(aes(size=weight)) +
  geom_smooth(method=lm, se=FALSE, color = "black") + # Add a regression line
  facet_wrap(~position) + # Break into individual graphs
  ggtitle("Relation of Weight to Height: All Positions") + # Title
  xlab("Height (In inches)") + ylab("Player's Weight") + # x and y labels
  theme_economist()
```
Q:7 - A historian would like to investigate the claim that the heights of 
players have increased over the years.  Analyze this claim graphically / visually. 
```{r Weight_Height_relate}
# Replace the value names
BBallP_C4 <- BBallP_C %>%
  mutate(height = recode(height, "6-0" = "72", "6-1" = "73", "6-2" = "74", 
                         "6-3" = "75", "6-4" = "76", "6-5" = "77", "6-6" = "78",
                         "6-7" = "79", "6-8" = "80", "6-9" = "81", "6-10" = "82",
                         "6-11" = "83", "7-0" = "84", "7-1" = "85", "7-2" = "86",
                         "7-3" = "87", "7-4" = "88", "7-5" = "89", "7-6" = "90", 
                         "7-7" = "91",)) %>%
        arrange(height, position )

# change the height column from chr to int
BBallP_C4 $height <- as.numeric(BBallP_C4$height)
BBallP_C4 

# Create a line plot using ggplot2
BBallP_C4 %>% 
ggplot(aes(x = year_start, y = height)) +
  geom_point() +
  geom_smooth(method=lm, se=FALSE, color = "red") + # Add a regression line
    facet_wrap(~position) +
  labs(title = "Heights of Players Since 1940",
       x = "Starting Year",
       y = "Height") +
  theme_economist()
```

Q:7 - (failed attempt) A historian would like to investigate the claim that the heights of 
players have increased over the years.  Analyze this claim graphically / visually.
```{r}
average_heights <- BBallP_C4 %>%
  group_by(year_start) %>%
  summarise(AverageHeight = mean(height))

# Create a line plot using ggplot2
BBallP_C4 %>% 
ggplot(aes(x = year_start, fill = height)) +
  geom_histogram() +
  facet_wrap(~position) +
  labs(title = "Trend in Player Heights Over the Years",
       x = "Year",
       y = "Height") +
  theme_minimal()

```

Q:8 - Create a 3D plot of height vs. weight vs. year and color code the 
points by position. 
```{r Weight_Height_relate}
# Replace the value names
BBallP_C4 <- BBallP_C %>%
  mutate(height = recode(height, "6-0" = "72", "6-1" = "73", "6-2" = "74", 
                         "6-3" = "75", "6-4" = "76", "6-5" = "77", "6-6" = "78",
                         "6-7" = "79", "6-8" = "80", "6-9" = "81", "6-10" = "82",
                         "6-11" = "83", "7-0" = "84", "7-1" = "85", "7-2" = "86",
                         "7-3" = "87", "7-4" = "88", "7-5" = "89", "7-6" = "90", 
                         "7-7" = "91",)) %>%
        arrange(height, position )

# change the height column from chr to int
BBallP_C4 $height <- as.numeric(BBallP_C4$height)
BBallP_C4 

# 3D plot 
p <- plot_ly(BBallP_C4, x = ~year_start, y = ~weight, z = ~height,
             color = ~position) %>%
   add_markers() %>%
   layout(title = "Heights of Players From 1940 to 2020", 
          scene = list(xaxis = list(title = 'Starting Year'),
                      yaxis = list(title = 'Weight'),
                      zaxis = list(title = 'Height')))
p
```

Q:9 - Go to this website and use one of the 50 best plots to visualize some 
aspect of the data and provide at least one insight.  You will present your 
work in breakout!
http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html

```{r factored boxplot}
g <- ggplot(BBallP_C4, aes(height, position))
g + geom_boxplot(aes(fill=factor(position))) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
  labs(title="Height of Players:", 
       subtitle="By Position",
       caption="Source: PlayersBBall",
       x="Height",
       y="Position")
```

```{r animate_year}
# Source: https://github.com/dgrtwo/gganimate
# install.packages("cowplot")  # a gganimate dependency
# devtools::install_github("dgrtwo/gganimate")
library(ggplot2)
library(gganimate)
library(gapminder)
theme_set(theme_bw())  # pre-set the bw theme.

p <- ggplot(
  BBallP_C4, 
  aes(x = height, y = weight, colour = position)
  ) +
  geom_jitter(show.legend = FALSE, alpha = 0.7) +
  scale_color_viridis_d() +
  scale_size(range = c(2, 12)) +
  scale_x_log10() + # convert to log scale
  labs(x = "Player Height", y = "Player Weight")
p
#p + transition_time(year_start) +
  #labs(title = "Year: {frame_time}")
p
#gganimate(g, interval=0.2)
```

# Data Import from csv
```{r csv_import3}
Path <- "D:/University/SMU/Doing_Data_Science/DDS_repository/Doing-Data-Science/Chapter2/Education_Income.csv"
Ed_Income = read.csv(Path, header=TRUE)

summary(Ed_Income)
unique(Ed_Income$Educ)
Ed_Income
```

```{r Education_Income_Relate}
# Replace the value names
Ed_Income3 <- Ed_Income %>%
  mutate(Educ = recode(Educ, ">16" = "16>", "<12" = "12-", "12" = "12-13")) %>%
        arrange(Educ, Income2005 )

Ed_Income2 <- Ed_Income
# change the height column from chr to int
Ed_Income2$Educ <- as.numeric(Ed_Income2$Educ)
Ed_Income2 

# Jitter plot with a line and labels in economist theme
Ed_Income2 %>% 
  ggplot(aes(x = Educ, y = Income2005)) + geom_jitter(aes(size=Income2005)) +
  geom_smooth(method=lm, se=FALSE) + # Add a regression line
  ggtitle("Relation of Education to total Earnings") + # Title
  xlab("Education Level") + ylab("2005 Income") + # x and y labels
  theme_economist()

# Use a boxplot to visualize the heights
Ed_Income3  %>% 
  ggplot(aes(x = Educ, y = Income2005)) + 
  geom_boxplot(color= "black", fill = "grey") +
  ggtitle("Relation of Education to total Earnings") + # Title
  xlab("Education Level") + ylab("2005 Income") + # x and y labels
  theme_economist()

```

# ChatGPT code: Question2 create a dataset to visually investigate the distribution of the 
weight of centers (C) is greater than the distribution of the weight of forwards (F). 
```{r chatGPT_scatter}

# Create a scatterplot using ggplot2
BBallC_F %>% 
  ggplot(aes(x = position, y = weight, color = position)) +
  geom_jitter() +
  labs(title = "Weight Distribution: Centers vs. Forwards",
       x = "Position",
       y = "Weight") +
  scale_color_manual(values = c("C" = "blue", "F" = "red")) +
  theme_economist()
```

# ChatGPT output
```{r CHatGPT_examples}
# Sample dataset (replace with your actual dataset)
height_data <- data.frame(
  Gender = c("Male", "Female", "Male", "Female", "Male", "Female"),
  Height = c(175, 162, 180, 157, 170, 165)
)

# Boxplot showing differences in height between men and women
ggplot(height_data, aes(x = Gender, y = Height, fill = Gender)) +
  geom_boxplot() +
  labs(title = "Differences in Height between Men and Women",
       x = "Gender",
       y = "Height") +
  scale_fill_manual(values = c("Male" = "blue", "Female" = "pink")) +
  theme_minimal()

# Density plot showing differences in height between men and women
ggplot(height_data, aes(x = Height, fill = Gender)) +
  geom_density(alpha = 0.5) +
  labs(title = "Differences in Height between Men and Women",
       x = "Height",
       y = "Density") +
  scale_fill_manual(values = c("Male" = "blue", "Female" = "pink")) +
  theme_minimal()
```




