---
title: "Unit 1"
author: "Adam E."
date: "2023-08-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r foundation, echo=FALSE}
install.packages("readr")
library(readr)

```

# Numeric dataframes
```{r age}
#Vectors
age = c(22,21,24,19,20,23)
age
age[2]
```

```{r years_math}
#Vectors
yrs_math_ed = c(4,5,2,5,3,5)
yrs_math_ed
```

```{r age_years_df}
# data frames
df1 = data.frame(Age = age, Years = yrs_math_ed)
df1
df1[4,2] # Data in the 4th row and 2nd column
df1[1, ] # all the data in the first row
df1$Years # all the data in the second column
```
# factor df - strings
```{r name_subject_vectors}
# vectors
a=c("Mary", "Martha", "Kim", "Kristen", "Amy", "Sam")
b=c("English", "Math", "Sociology", "Math", "Music", "Dance")
a
b
```

```{r name_subject_df}
# vectors
df2 = data.frame(Name=a, Major=b)
df2
```

```{r four_column_df}
# vectors
df3 = data.frame(Age=age, Years=yrs_math_ed, Name=a, Major=b)
df3
```

#Cbind combines columns
```{r bind_two_dfs}
# combine two dfs
df4 = cbind(df1,df2)
df4
```

#add a new row
```{r insert_row_in_df}
# add a row
d=c(19,4,"John","Math")
d
# Add a new row
df5 = rbind(df4,d)
df5
summary(df5)
```

# cast a value as a character
```{r cast_character_df}
dfCopy = df4
# Cast as a character
dfCopy$Name = as.character(df4$Name)
dfCopy$Major = as.character(df4$Major)
summary(dfCopy)

```

#add a new row
```{r insert_row_in_df}
# add a row
dfCopy2 = rbind(dfCopy,d)
dfCopy2
summary(dfCopy2)
```
# fix the classes
```{r cast_numeric/factor_df}
# Cast as a numeric
dfCopy2$Age = as.numeric(dfCopy2$Age)
dfCopy2$Years = as.numeric(dfCopy2$Years)
summary(dfCopy2)
# Cast as a factor
dfCopy2$Name = as.factor(dfCopy2$Name)
dfCopy2$Major = as.factor(dfCopy2$Major)
summary(dfCopy2)
```

# filter the df
```{r cast_numeric/character_df}
# All students with more than 4 years of Math
df5[df5$Years > 4,]
# All students with more than 4 years of Math and are 21 years of age or older.
df5[(df5$Age >= 21 & df5$Years > 4), ]
# All students that are majoring in Math
df5[df5$Major == "Math",]

```

# Data Import from csv
```{r csv_import}
Path <- "D:/University/SMU/Doing_Data_Science/DDS_repository/Doing-Data-Science/Chapter1/Creativity.csv"
Example1 = read.csv(Path, header=TRUE)
Example1

```

# Data Import from csv
```{r csv_import2}
Path <- "D:/University/SMU/Doing_Data_Science/DDS_repository/Doing-Data-Science/Chapter1/BusinessSales.csv"
Example2 = read.csv(Path, header=TRUE)
Example2
# file.choose() - pops up behind Rstudio.
#Example3 = read.csv(file.choose()) 
#Example3

```

### Part 2: Visualization
# Scatterplots
# Plot(x,y,col,type,ylab,xlab,main)
```{r csv_plotting}
# plot(x-axis data, y-axis data, pch =plot character, xlab=x label, ylab=y label, main= main title label)
plot(Example2$ad_tv, Example2$sales, pch=15,xlab="TV Advertising $$$", ylab="Sales $$$", main="Sales v. Advertising")
# add a line (h=horizontal line, col=color, lwd=line width )
abline(h=55, col = "red", lwd= 5)

```
### plotting wth mpg - ggplot2 required
```{r}
plot(mpg$hwy,mpg$cty,pch = 15) #NO LABLES ... AHHH!!!
plot(mpg$hwy,mpg$cty,pch = 15, main = "City MPG v. Highway MPG", ylab = "City MPG", xlab = "Highways MPG")
```
### plotting wth iris - Petal Length vs. Sepal Length
```{r iris_data}
plot(iris$Sepal.Length, iris$Petal.Length, pch = 15) #NO LABLES ... AHHH!!!
plot(iris$Sepal.Length, iris$Petal.Length, pch = 15, main = "Iris Flower Parts comparison", ylab = "Petal Length", xlab = "Sepal Length")
```
### plot only viginica
```{r}
# Subset the species column where the value == virginica
irisVir = iris[iris$Species == "virginica",]
# plot this subset in blue. Note: define the vector location of the labels to limit the plot size.
plot(irisVir$Sepal.Length,irisVir$Petal.Length, col = "blue", ylim = c(0,7), xlim = c(4,8))
# Subset the species column where the value == virsicolor
irisVers = iris[iris$Species == "versicolor",]
# use points vice plot to add point to the existing plot rather than a new plot.
points(irisVers$Sepal.Length,irisVers$Petal.Length, col = "red")
# Subset the species column where the value == setosa
irisSet = iris[iris$Species == "setosa",]
points(irisSet$Sepal.Length,irisSet$Petal.Length, col = "green")

#Note: when just using a regular r file, you can use dev.off() to clear the plot and start fresh.
```
### Dividing the plot space!  c(num rows, num columns)
```{r}
# Function par(mfrow = c(number of columns, number of rows)) divides the plot space.
par(mfrow = c(1,2))
# histograms.
hist(mpg$cty,col = "blue", main = "Histogram of MPG", xlab = "MPG")
# Boxplots
boxplot(mpg$cty~mpg$class, data= mpg, main = "Boxplot of MPG by Class", xlab = "Cylinder")
```


### Dividing the plot space!  c(num rows, num columns)
# Iris sepal lengths
```{r}
# Function par(mfrow = c(number of columns, number of rows)) divides the plot space.
par(mfrow = c(1,2))
# histograms.
hist(irisSet$Sepal.Length,col = "blue", main = "Histogram of Iris Sepal Lengths", xlab = "Sepal")
# Boxplots
boxplot(irisSet$Sepal.Length~irisSet$Petal.Length, data= iris, main = "Boxplot of iris by Class", xlab = "Sepal", ylab = "Petal")
```
### barplot
```{r}
# age vectors
age = c(22,21,24,19,20,23)
# years of math vectors
yrs_math_ed = c(4,5,2,5,3,5)
# define the names and subjects
names = c("Mary","Martha","Kim","Kristen","Amy","Sam")
subject = c("English","Math","Sociology","Math","Music","Dance")
# Create the data frame
df3 = data.frame(Age = age, Years = yrs_math_ed, Name = names, Subject = subject)
# ploat as a barplot with years on the y axis and the x axis divided by names
barplot(df3$Years, names.arg = df3$Name, ylab="Math Years", main="Math Years")
```
### We want a barplot of frequency of each class of car.
###note here how the data is in a different format for mpg.
###it is not one row per relevant observation with the value at the end
###we need to actually count the number of each class and then plot

### We could count ... that would take a while

### We could turn class into a factor and then let summary count for us.
```{r}
#
summary(mpg$class)
# Cast the data in class as a factor and assign it to a new column called classFact.
mpg$classFact = as.factor(mpg$class)
# Review the head of the data.
head(mpg)
# Verify the new column is a factor.
summary(mpg$classFact)
# Plot the summary.
barplot(summary(mpg$classFact), main="Number of Auto Classes", ylab="Count")
```
```{r}
# Function pairs() creates a series of scatter plot pairs.
pairs(~mpg+disp+drat+wt,data=mtcars, main="Simple Scatterplot Matrix")
```

# UNIT 1 R Review Part 3.R
## Statistics in R

### Sampling Distribution of the Individual
```{r}
#draw a sample from a standard normal distribution
#run many times varying sample size and look at histogram and mean
#Random normal Parameters: rnorm(sample, mean, standard deviation).
sample1 = rnorm(1000,0,1)
hist(sample1)
mean(sample1)
sd(sample1)
```

### Another way to code the same thing
```{r}
population = rnorm(10000000,0,1) #note the the number of draws here
hist(population)
# Function: sample(dataset, draw count)
sample1 = sample(population,100) #sample of size 100
hist(sample1)
mean(sample1)
sd(sample1)
```

### Sampling Distribution of the Sample Mean
# now make a function in which you pass the sample size, the number of samples to take #and thus the number of sample means to return as well as the mean and sd.  

# Define these outside the function so that they exist outside the function (global)

```{r}
xBarVec = c() #Global vector to hold the sample means
#Random normal Parameters: rnorm(sample, mean, standard deviation).
population = rnorm(10000000,0,1) #Simulating the population
```

### The Function
```{r}
#####################################################
# Function: xbarGenerator
# Arguments: samplesize: the size of the sample that each sample mean is based on. 
#            number_of_samples: the number of samples and thus sample means we will generate
# Author: Bivin Sadler
#####################################################
# Use function() to create a function and assign it to xbarGenerator
xbarGenerator = function(sampleSize = 30, number_of_samples = 100)
{
  for(i in 1:number_of_samples) # 1:n is like looking at sample 1 up to n or 100
  {
    theSample = sample(population,sampleSize) # Function: sample(dataset, draw count)
    xbar = mean(theSample)
    xBarVec = c(xBarVec, xbar)
  }
  return(xBarVec)
}
```

### Before you Run This.... What do you think will be the "Sampleing Distribution of the Sample Mean"?
```{r}
xbars = xbarGenerator(30,100)
length(xbars)
hist(xbars)
```

## You Try It
Now adapt the code above to generate a specified number of sample means of a specified size sample size (this has already been done) from a normally distributed population with a specified mean and standard deviation (currently it is forced to be from a standard normal distribution N(0,1)).

Run it a few times and use it generate 10000 sample means, each from a sample of size 50, from a normal distribution with mean = 60 and sd = 10.  Provide a distribution (histogram) of the distribtions as well. 

What is the shape of the distribution?  What is the mean and sd?  What should they be given the Central Limit Theorem (CLT).  For a refresher of the CLT see the Stats Bridge Course!


### My attempted Function
```{r}
#####################################################
# Function: xbarGenerator
# Arguments: samplesize: the size of the sample that each sample mean is based on. 
#            number_of_samples: the number of samples and thus sample means we will generate
# Author: Bivin Sadler
#####################################################

# Use function() to create a function and assign it to xbarGenerator
xbarGenerator2 = function(sampleSize = 30, number_of_samples = 100, sample_mean = 0, sample_sd = 1)
{
  population = rnorm(10000000,sample_mean, sample_sd) #Simulating the population
  
  for(i in 1:number_of_samples) # 1:n is like looking at sample 1 up to n or 100
  {
    theSample = sample(population,sampleSize) # Function: sample(dataset, draw count)
    xbar = mean(theSample)
    xBarVec = c(xBarVec, xbar)
  }
  return(xBarVec)
}
```

### Before you Run This.... What do you think will be the "Sampleing Distribution of the Sample Mean"?
```{r}
xbars2 = xbarGenerator2(50,10000,60,10)
length(xbars2)
hist(xbars2)
summary(xbars2)
sd(xbars2)
10/sqrt(60)
```












### Possibe Solution
```{r}
xBarVec = c() #global vector to hold the sample means
#####################################################
# Funciton: xbarGenerator (Adpated)
# Argements: samplesize: the size of the sample that each sample mean is based on. 
#            number_of_samples: the number of samples and thus sample means we will generate
# Author: Bivin Sadler
#####################################################
xbarGenerator2 = function(sampleSize = 30,number_of_samples = 100, mean = 0, sd = 1)
{
  for(i in 1:number_of_samples)
  {
    theSample = rnorm(sampleSize,mean,sd)
    xbar = mean(theSample)
    xBarVec = c(xBarVec, xbar)
  }
  return(xBarVec)
}

xbars = xbarGenerator2(50,10000,60,10)
hist(xbars)
summary(xbars)
sd(xbars)
10/sqrt(60)
```




